\bfseries A simple momentum-transport model for sailing

### 2021-07-06 paper/main.tex
Tao re-read; GoodSailing(tm)

% to-do items:
% ------------
% - FIND THE PRIOR LITERATURE!!
% - Should Terry Tao be cited way earlier?
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.

\documentclass[letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry} % because Overleaf is weird.
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{./}{../ipynb/}}
% \usepackage{biblatex}
% \addbibresource{sailing.bib}

@@

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\textheight}{9.00in}
\setlength{\topmargin}{-0.40in}
\pagestyle{myheadings}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

@@

Our model for forces on a sailboat presented below in \secref{sec:model} is a very generic momentum-transport model.
In astrophysics contexts, forces from momentum transport from winds or relative motions of bodies and fluids are called ``ram-pressure'' forces.
We use that term a bit in what follows, but since it means different things to different people, it's fine to substitute in the word ``lift'' for ``ram-pressure'' everywhere.
We'll discuss these forces conceptually in \secref{sec:ram}, give Galilean-relativistic expressions in \secref{sec:model}, and use them to explain the steady-state motion of a sailboat in \secref{sec:steady}.
We follow this modeling with some optimization, useful to an idealized sailor on our idealized boat, in \secref{sec:good} and \secref{sec:better}.
When we optimize the trim of our boat, we learn about what things the boat can do, in terms of sailing speed, and sailing upwind and downwind.
It's impressive!
We will discuss the dependence of our results on our boat properties in \secref{sec:design} and the limitations and extensions of our model in \secref{sec:discussion}.

@@

In \secref{sec:steady} we showed how the boat moves (in steady state), given any arbitrary setting of the orientation of the keel and the sail.
However, one doesn't sail by randomly setting the keel and sail!
One sails by pointing the boat in some direction and then setting the sail (or sails) appropriately.
In our simple, ram-pressure boat, how should we set the planar sail?

HOGG: IN THIS SECTION EXPLICITLY DEFINE ``GOOD SAILING''.
\begin{align}\label{eq:good}
    \uvec_{\perp\sail}^\good &\leftarrow \argmax_{\uvec_{\perp\sail}} \left[\uvec_{\parallel\keel}^\top\,(\vboat-\vwater)\right] ~,
\end{align}
where .... HOGG IMPLICITLY $\vboat-\vwater$ is a function of all the properties of the world and the boat boat, along with the orientation of its keel and sail.

HOGG: FIGURE showing good sailing for a set of boat orientations.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  GoodSailing(tm) is unsuccessful in the last two cases, where the pointing is too close to directly upwind.
  In these cases the the maximum of the GoodSailing(tm) objective \eqref{eq:good} is negative.\label{fig:good}}
  \figurerule
\end{figure}

HOGG: Announce and discuss \figref{fig:hourgood}.
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set of randomly oriented keels.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.
  As is visible in \figref{fig:good}, when the keel orientation is too close to directly upwind, the boat goes downwind instead; there are no \emph{directly} upwind options.\label{fig:hourgood}}
  \figurerule
\end{figure}

HOGG: Figure showing direction cosines of sail relative to keel/boat orientation.

HOGG: Comment: The settings of the sail look ``tight'' relative to standard sailing practice. But this boat has flat sails, so it's hard to compare the exact settings with sailor experiences (which are all with curved sails).

HOGG: Final comment: But this is not the BEST a sailor can do...

@@

% \raggedright
\footnotesize
\bibliographystyle{annotate}
\bibliography{sailing.bib}
% \printbibliography
\end{document}

>>>>

% to-do items:
% ------------
% - Change citations to Tao to make it clear that the model is wrong!!
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.

\documentclass[letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper]{geometry} % because Overleaf is weird.
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{./}{../ipynb/}}
% \usepackage{biblatex}
% \addbibresource{sailing.bib}

@@

% typesetting adjustments
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
  {-3.25ex \@plus -1ex \@minus -.2ex}%
  {1.5ex \@plus .2ex}%
  {\raggedright\normalfont\large\bfseries}}
\makeatother
\setlength{\textwidth}{5.00in}
\setlength{\oddsidemargin}{3.25in}
\addtolength{\oddsidemargin}{-0.5\textwidth}
\setlength{\textheight}{9.00in}
\setlength{\topmargin}{-0.40in}
\pagestyle{myheadings}
\newcommand{\figurerule}{\rule[1ex]{\textwidth}{0.2pt}}
\linespread{1.08}
\frenchspacing\sloppy\sloppypar\raggedbottom

@@

Our model for forces on a sailboat presented below in \secref{sec:model} is a very generic momentum-transport model.
In astrophysics contexts, forces from momentum transport from winds or relative motions of bodies and fluids are called ``ram-pressure'' forces.
We use that term a bit in what follows, but since it means different things to different people, it's fine to substitute in the word ``lift'' for ``ram-pressure'' everywhere.
We'll discuss these forces conceptually in \secref{sec:ram}, give Galilean-relativistic expressions in \secref{sec:model}, and use them to explain the steady-state motion of a sailboat in \secref{sec:steady}.
We follow this modeling with some optimization, useful to an idealized sailor on our idealized boat, in \secref{sec:good} and \secref{sec:better}.
When we optimize the trim of our boat, we learn about what things the boat can do, in terms of sailing speed, and sailing upwind and downwind.
It's impressive!
We will discuss the dependence of our results on our boat properties in \secref{sec:design}.
We comment on the implemntation of the model in \secref{sec:implementation} and we remark on limitations and extensions in \secref{sec:discussion}.

@@

In \secref{sec:steady} we showed how the boat moves (in steady state), given any arbitrary setting of the orientation of the keel and the sail.
However, one doesn't sail by randomly setting the keel and sail!
One sails by pointing the boat in some direction and setting the sail (or sails) appropriately given that boat pointing, or vice versa.
In our simple, ram-pressure boat, how should we set the planar sail?

We could define \emph{GoodSailing(tm)} to be setting the sail orientation optimally, given the orientation of the boat.
This might look something like the following:
\begin{align}\label{eq:good}
    \uvec_{\perp\sail}^\good &\leftarrow \argmax_{\uvec_{\perp\sail}} \left[\uvec_{\parallel\keel}^\top\,(\vboat-\vwater)\right] ~,
\end{align}
where $\uvec_{\perp\sail}^\good$ is the best orientation of the (planar) sail conditioned on the orientation of the keel,
$\uvec_{\parallel\keel}$ is the unit vector pointing along the keel (parallel to the keel, in the forwards direction),
the inner product gives the projection of the boat--water velocity vector onto the boat heading direction,
and (for optimization purposes) the velocity difference $\vboat-\vwater$ is implicitly a function of the sail and keel orientations, according to the model of \secref{sec:steady}.
This optimization is one-dimensional, but it is a global optimization, so it requires some attention; details of our implmentation of the optimization are given in \secref{sec:implementation}.
Some examples of GoodSailing(tm) settings of sail and keel are shown in \figref{fig:good}, along with the resulting boat velocity vectors.
A summary of all possible boat--water velocity vectors possible, for GoodSailing(tm) at many different keel orientations, is shown in \figref{fig:hourgood}.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  GoodSailing(tm) is unsuccessful in the last two cases, where the pointing is too close to directly upwind.
  In these cases the the maximum of the GoodSailing(tm) objective \eqref{eq:good} is negative.\label{fig:good}}
  \figurerule
\end{figure}
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set of randomly oriented keels.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.
  As is visible in \figref{fig:good}, when the keel orientation is too close to directly upwind, the boat goes downwind instead; there are no \emph{directly} upwind options. This figure is directly comparable to [HOGG WHAT FIGURES IN THE BOOKS ABOUT SAILING?].\label{fig:hourgood}}
  \figurerule
\end{figure}

HOGG: Figure showing direction cosines of sail relative to keel/boat orientation?

HOGG: Comments on GoodSailing(tm): As we saw in \secref{sec:steady}, you can sail dwfttw, and you can sail upwind, and you can sail fttw in general. But you can't sail dufttw or ddfttw. The settings of the sail look ``tight'' relative to standard sailing practice. But this boat has flat sails, so it's hard to compare the exact settings with sailor experiences (which are all with curved sails).

HOGG: Final comment: But this is not the BEST a sailor can do, because (a) boat doesn't move in the $\uvec_{\parallel\keel}$ direction, (b) because the destination might also be moving wrt the water, and (c) because you might want to take a tacking path...

@@

% \raggedright
\bibliographystyle{hogg-plain-annote}
\bibliography{sailing.bib}
% \printbibliography
\end{document}
